---
title: "quant project"
output: html_document
---

Load required packages
```{r}
library(ggplot2)
library(permute)
library(lattice)
library(vegan)
library(pwr)
library(grid)
library(gridExtra)
```


Randomly generate data for all of the parameters I will measure for my study
```{r}
tank<-sample(c("S1", "S2", "S3", "S4", "ENV"), 183, replace = TRUE)
sdens<-as.numeric(sample(seq(1000000,9000000, 100000), 183, replace=TRUE))
swim<-as.numeric(sample(seq(0.5, 7, 0.01), 183, replace=TRUE))
prHb<-sample(seq(5.9,6.3, 0.01), 183, replace=TRUE)
poHb<-sample(seq(6.2,6.9, 0.01), 183, replace=TRUE)
prHct<-sample(seq(25,28, 0.1), 183, replace=TRUE)
poHct<-sample(seq(27,31, 0.1), 183, replace=TRUE)
bLac<-sample(seq(4.0,6.8, 0.01), 183, replace=TRUE)
bGlu<-sample(seq(4.6,8, 0.01), 183, replace=TRUE)
mLac<-sample(seq(24,30, 0.01), 183, replace=TRUE)
mGlu<-sample(seq(1.5,3.3, 0.01), 183, replace=TRUE)
mGly<-sample(seq(1.5,3.8, 0.01), 183, replace=TRUE)
CS<-sample(seq(0.1,0.5, 0.01), 183, replace=TRUE)
LDH<-sample(seq(0.2,0.8, 0.01), 183, replace=TRUE)
leng<-sample(119:200, 183, replace=TRUE)
sex<-sample(c("M", "F"), 183, replace = TRUE)
```

Create a dataset using those randomly generated values
```{r}
SSTdata<-data.frame(tank, sdens, swim, prHb, poHb, prHct, poHct, bLac, bGlu, mLac, mGlu, mGly, CS, LDH, leng, sex)

```

```{r}
print(SSTdata)
```



 Create a linear model to see relationship between myxospore density and swim time

```{r}
timexspore<-lm(as.numeric(SSTdata$swim) ~ as.numeric(SSTdata$sdens))
summary(timexspore)
```

Plot lm from above
```{r}
ggplot(SSTdata, aes(x = as.numeric(sdens), y = as.numeric(swim))) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  ggtitle("Relationship Between Swim Time and Myxospore Density") +
  labs(y= "swim time to exhaustion (min)", x = "myxospore density")
```

Define physio to include all physiological parameters including hematocrit/ hemoglobin/ lactate/ glucose/ glycogen/ CS/ LDH

run a lm to see how spore density combined with physiological parameters affects swim time
```{r}
physio<-as.numeric(SSTdata$prHb)+as.numeric(SSTdata$poHb)+as.numeric(SSTdata$prHct)+as.numeric(SSTdata$poHct)+as.numeric(SSTdata$bLac)+as.numeric(SSTdata$mLac)+as.numeric(SSTdata$bGlu)+as.numeric(SSTdata$mGlu)+as.numeric(SSTdata$mGly)+as.numeric(SSTdata$CS)+as.numeric(SSTdata$LDH)

ts2<-glm(as.numeric(SSTdata$swim) ~ as.numeric(SSTdata$sdens) + physio)

summary(ts2)
```

plot lm from above
```{r}
ggplot(SSTdata, aes(x = as.numeric(sdens) + physio, y = as.numeric(swim))) + 
  geom_point() +
  stat_smooth(method = "glm", col = "red") +
  ggtitle("Relationship Between Swim Time and Physiological Parameters") +
  labs(y= "swim time to exhaustion (min)", x = "physiological suite")
```

run anova and tukey to look for differences between 5 different tanks
```{r}
tankaov<-aov(as.numeric(sdens) ~ tank, data=SSTdata)
summary(tankaov)
```

```{r}
tanktukey<-TukeyHSD(tankaov)
tanktukey
```

create a subset of SST that excludes non-numeric columns (tank and sex)
```{r}
names(SSTdata) %in% c('tank', 'sex')
subSST<-SSTdata[, !(names(SSTdata) %in% c('tank', 'sex'))]
subSST
```

PCA and plot of numeric SST subset
```{r}
PCA1<-rda(subSST ~ 1, scale = TRUE)
```


```{r}
plot(PCA1, main="Seatrout Data PCA")

```

```{r}
summary(PCA1)
```

compute what amount of variance is explained by each PCA axis
```{r}
PCA1$
PCA1$tot.chi
PCA1$CA$eig
```
```{r}
round(PCA1$CA$eig / PCA1$tot.chi, 2)
```


create new subset including all parameters of physiological nature, as well as swim time (expected to be connected in real exp)
```{r}
physSST<-SSTdata[, !(names(SSTdata) %in% c('tank', 'sex', 'leng', 'sdens'))]
physSST
```

RDA investigating effects of myxospore density on physiological suite
```{r}
PCA2<-rda(physSST ~ SSTdata$sdens, scale = TRUE)
```

```{r}
summary(PCA2)
```

```{r}
plot(PCA2, main="Seatrout Data RDA")
```

```{r}
PCA2$
PCA2$tot.chi
PCA2$CA$eig
```
```{r}
round(PCA2$CA$eig / PCA1$tot.chi, 2)
```




power analysis to see how many fish needed per tank to ensure high power
```{r}
p1<-pwr.anova.test(f=0.25,n=50,k=5,sig.level=0.05)
p1
```

convertSSTdata to a CSV file
```{r}
write.csv(SSTdata,"C:\\Users\\Jessica Daly\\Documents\\R\\SST-project\\SSTdata.csv", row.names = TRUE)
```

attempt at setting up PCA2 in ggplot
```{r}
PCA2_out <- as.data.frame(PCA2$x)
PCA2_out$group <- sapply( strsplit(as.character(row.names(df)), "_"), "[[", 1 )
head(PCA2_out)
```

```{r}
ggPCA2<-ggplot(PCA2_out,aes(x=SSTdata$sdens,y=physio,color=tank)) +
        ggtitle("Relationship Between Myxospore Density and Physiological Parameters") +
        labs(y= "physiological suite", x = "myxospore density")
ggPCA2<-ggPCA2+geom_point()
ggPCA2
```







